---
title: "Halloween Candy"
author: "Stephanie Duncan"
date: "16/12/2020"
output: html_document
---

```{r}
#Loading the relevent libraries
library(tidyverse)
library(dplyr)
library(readxl)
library(janitor)
library(tidyr)
library(stringr)
library(assertr)
library(here)
library(readr)
```

```{r}
#Testing where the top level of the project directory is
here::here()

#Setting the path to the data file
halloween_candy <- read_csv(here("clean_data/halloween_candy.csv"),
#Parsing failures error solution for gender and country columns
col_types = cols(gender = col_character(), country = col_character()))
```

1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
total_ratings <-
halloween_candy %>% 
  filter(!is.na(rating)) %>% 
  summarise(total_ratings = n())

#There are 548814 total ratings.

```

2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
average_age_trick_or_treating <- halloween_candy %>% 
  group_by(trick_or_treating) %>%
 drop_na() %>% 
  summarise(avg_age = round(mean(age, na.rm = TRUE)))
#The average age of people who are trick or treating is 39 compared to an average age of 42 for those not trick or treating.
```
| trick_or_treating | average_age |
|-------------------|-------------|
| No                | 42          |
| Yes               | 39          |


3. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
#Find the ratings total for each bar for each rating
most_rated_candy <- halloween_candy %>% 
   select(candy, rating) %>% 
  filter(rating == c("joy", "despair", "meh")) %>% 
  group_by(candy, rating) %>% 
  count() 

#Show the ratings 
most_rated_candy <- most_rated_candy %>% 
  group_by(rating) %>% 
  slice_max(n) %>% 
  arrange(desc(n))
```
| candy                      | rating  | ratings_count |
|----------------------------|---------|---------------|
| reese_s_peanut_butter_cups | joy     | 3307          |
| gum_from_baseball_cards    | despair | 3265          |
| chiclets                   | meh     | 879           |

4. How many people rated Starburst as despair?

```{r}
starbust_despair_rating <- halloween_candy %>% 
  filter(candy == "starburst",
         rating == "despair") %>% 
  summarise(total = n())
#1865 people
```

For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
halloween_candy <- halloween_candy %>% 
  filter(rating == c("despair", "joy", "meh")) %>% 
  mutate(rating = case_when(
    rating == "despair" ~ -1, 
    rating == "joy" ~ +1, 
    rating == "meh" ~ 0
  ))
```


5. What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
popular_candy_by_gender <- halloween_candy %>% 
  select(gender, candy, rating) %>% 
  drop_na() %>% 
  group_by(gender, candy) %>% 
  summarise(total = sum(rating)) %>% 
  slice_max(total) %>% 
  arrange(desc(total))
```

| gender             | candy | total |
|--------------------|-------|-------|
| male               | twix  | 992   |
| female             | twix  | 574   |
| I'd rather not say | twix  | 41    |
| other              | twix  | 24    |

6. What was the most popular candy bar in each year?

```{r}
most_popular_candy_by_year <-
halloween_candy %>% 
  select(year, rating, candy) %>% 
  drop_na() %>% 
  group_by(year, candy) %>% 
  summarise(total = sum(rating)) %>% 
  slice_max(total)
```

| year | candy                    | total |
|------|--------------------------|-------|
| 2015 | any_full_sized_candy_bar | 1458  |
| 2016 | any_full_sized_candy_bar | 322   |
| 2017 | twix                     | 1376  |


7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
most_popular_candy_by_countries <-
halloween_candy %>% 
  drop_na(country) %>% 
  select(country, rating, candy) %>% 
    mutate(country = case_when(
      country == "usa" ~ "usa",
      country == "canada" ~ "canada",
      country == "uk" ~ "uk",
      TRUE ~ "other"
    )) %>% 
  group_by(country, candy) %>% 
  summarise(total = sum(rating)) %>% 
  slice_max(total) %>% 
  arrange(desc(total))
```

| country | candy        | total |
|---------|--------------|-------|
| usa     | twix         | 1437  |
| canada  | mars         | 150   |
| other   | butterfinger | 31    |
| other   | twix         | 31    |
| uk      | twix         | 24    |
